---
title: "STAT545_hw005.Rmd"
output: html_document
---

# Introduction
This is Zachary Sherker's hw005 for STAT 545.

In this assignment, I will:

1)Reorder a factor based on the data and demonstrate the effect in arranged data and in figures.
2)Write data to file and load it back into R.
3)Improve a figure through the use of factor levels, smoother mechanics, color schemes.
4)Convert this to a plotly visual.
# Part 1: Factor management
First, I will start by uploading the required tools and data sets
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(plotly))  
suppressPackageStartupMessages(library(scales))
```
## Elaborating the `gapminder` data set
(1) Drop Oceania
I will begin by filtering Oceania data out of the `gapminder` data set:
```{r}
Oceania_drop <- gapminder %>% 
  filter(continent != "Oceania")
str(Oceania_drop)
```
I will then drop the unused Oceania data:
```{r}
First_drop <- Oceania_drop %>% 
  mutate(continent=fct_drop(continent))
  str(First_drop)
```
```{r}
Second_drop <- Oceania_drop %>% 
  droplevels() 
  str(Second_drop) 
```
(2) Reorder the levels of `country` or `continent`
The conintents in the `gapminder` dataset are currently ordered alphabetically:
```{r}
levels(gapminder$continent)
```
We will first reorder the continents by mean GDPpercapita (mean GDP shown in plot):
```{r}
gapminder %>% 
  mutate(continent=fct_reorder(continent,gdpPercap,mean)) %>% 
ggplot(aes(continent,gdpPercap)) + geom_violin(aes(fill=continent))+
  stat_summary( fun.y=mean, colour="green", geom="point", size=2,show.legend  = TRUE ) +
  stat_summary( fun.y=mean, colour="purple", geom="text", size = 4, show.legend  = TRUE, 
               vjust=-0.7, aes( label=round( ..y.., digits=1 ) ) ) 
```
I will now reorder the continents by maximum gdpPercapita (max GDP shown in plot):
```{r}
gapminder %>% 
  mutate(continent=fct_reorder(continent,gdpPercap,max)) %>% 
ggplot(aes(continent,gdpPercap)) + geom_violin(aes(fill=continent))+
stat_summary( fun.y=max, colour="green", geom="point", size=2,show.legend  = TRUE ) +
  stat_summary( fun.y=max, colour="purple", geom="text", size = 4, show.legend  = TRUE, 
               vjust=-0.7, aes( label=round( ..y.., digits=1 ) ) ) 
```
Finally, I will reorder the continents by minimum GDPpercapita (min. GDP shown in plot):
```{r}
gapminder %>% 
  mutate(continent=fct_reorder(continent,gdpPercap,min)) %>% 
ggplot(aes(continent,gdpPercap)) + geom_violin(aes(fill=continent))+
  stat_summary( fun.y=min, colour="green", geom="point", size=2,show.legend  = TRUE ) +
  stat_summary( fun.y=min, colour="purple", geom="text", size = 4, show.legend  = TRUE, 
               vjust=-0.7, aes( label=round( ..y.., digits=1 ) ) ) 
```
## Part II: File I/O
I start by filtering the data to only show information from the Americas in 2002:
```{r}
filterdata <- gapminder %>% 
  filter(continent == "Americas" & year == 2002) 

# drop unused levels
AmericasData <- filterdata %>% 
 droplevels() 
# check the levels of continent and country to be sure unused data is dropped.
str(AmericasData)

```
## write/read csv
I will now write the new dataset out in a csv file:
```{r}
write_csv(AmericasData,"AmericasData.csv")
```
And read it back in as a csv file:
```{r}
read_AmeicasData<- read_csv("AmericasData.csv")
```
I will now check the newly read-in dataset:
```{r}
head(read_AmeicasData)
```
## save/read rds
```{r}
# save to RDS file
saveRDS(AmericasData, "AmericasData.rds")

# read from RDS file
read_RDSdata <- readRDS("AmericasData.rds")

# check readin data
 head(read_RDSdata) 
```
## Put data into text file and subsequently read from text file
```{r}
# put data into text file
dput(AmericasData, "AmericasData.txt")

# retrieve data from text file
data_txt <- dget("AmericasData.txt")
 
head(data_txt) 
```
## Part III: Vizualization of data
I will start by creating a basic graph comparing the GDPpercap of all countries within continental groupings:
```{r}
ggplot(gapminder,aes(gdpPercap,continent))+
  geom_line(aes(colour=continent,size=gdpPercap),alpha=0.8)
```
I will now modify the graph to make it more informative by first reorganizing the data:
```{r}
# get max, min, median and mean GDP for each contient in all years
  Reorganized_data <-  gapminder %>% 
  group_by(continent,year) %>% 
summarize(
  min_gdp = min(min(gdpPercap)),
  max_gdp = max(max(gdpPercap)),
  mean_gdp = mean(mean(gdpPercap))
)
Reorganized_table <- gather(Reorganized_data,key = "Type_GDP", value="Value_GDP", min_gdp, max_gdp,mean_gdp)
# then check the new gathered table
knitr::kable(head(Reorganized_table))
```
Now I will plot this data:
```{r}
Reorganized_graph <-  Reorganized_table %>% 
  ggplot(aes(x = year, y = Value_GDP, color = Type_GDP) ) +
  facet_wrap(~continent) +
  scale_y_log10(label=dollar_format())+
  scale_x_continuous()+
  geom_point()+
  geom_line()+
       labs(x = "year",
          y = "GDP",
          title = "Variables of GDPpercap per continent per year")+
  theme(axis.text = element_text(size= 10),
          strip.background = element_rect(fill = "green"),panel.background = element_rect(fill = "white"))
  
  Reorganized_graph
```
As you can see, this reorganized data makes for a much more informative graph, allowing us to observe larger trends in a simple format.
(2) Convert graph to `plotly`
```{r}
ggplotly(Reorganized_graph)
```
The plotly version of this graph allows for us to access the information portrayed by simply scrolling our mouse over the data points, making the graph much more informative still.
## Part IV: Writing figures to file
```{r}
ggsave("modified_graph.png", width=16, height=6, units = "cm")
```

